The eclipse of ASASSN-21qj

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

The star 2MASS J08152329-3859234 underwent a sudden dimming event in December 2021, which was announced by \citet{RizzoSmith21} and assigned the identifier \asas, which we will subsequently use for the rest of this paper.
%
The star has undergone rapidly fluctuating photometry through to August 2022 \citep{RizzoSmith22} and has been monitored intensively by the AAVSO observers and the LCOGT network of telescopes.
%
It had previously shown no significant stellar variation in the optical bands for 2300 days as reported in \citet{RizzoSmith21}.
%
Searches through other photometric archives showed no other significant changes in the optical bands before this epoch.
%
The wide field infrared satellite WISE has photometric imaging from 3.8 microns (band W1) through to 22 microns (band W4), and the NEOWISE survey has photometry for bands W1 and W2 for this star.
%
This star showed a significant brightening of 0.7 magnitudes in W1 and 0.8 magnitudes in W2 between two observing epochs (65000 MJD and 65120 MJD), and the IR color of the star had changed from W1-W2=0.0 to 1.2, but no significant changes in flux in the optical bands were seen during this time.
%
Some 900 days later, the dimming was seen in the optical, and the absorption is larger at shorter optical wavelengths.
%
These observations are all consistent with an event that generated a significant amount of sub-micron dust which subsequently started to transit the stellar disk.

@@

The properties of \asas\ (Gaia EDR3 source 5539970601632026752) are listed in Table~\ref{tab:Stellarprop}.
%
Photometry from various sources was compiled and analysed using the Virtual Observatory SED Analyzer \citep[VOSA; ][]{Bayo08} and the results are listed in Table~\ref{tab:Stellarprop}, showing that \asas\ is consistent with being a G2 type dwarf star.

@@

\subsection{Possible multiplicity of \asas}

\asas\ (Gaia DR3 5539970601632026752 = 2MASS~J08152329-3859234) has a neighbor (Gaia DR3 5539970597334497024 = 2MASS~J08152298-3859244) which might plausibly be a physical companion.
%
Based on the Gaia EDR3 mean ICRS position for epoch 2016.0, the visual companion lies at a separation $\rho$ = $3738.243\pm0.062$ mas and at position angle $\theta$ = $249^{\circ}.977$.
%
Their parallaxes ($\varpi$ = $1.7631\pm0.0112$ mas vs. $1.4711\pm0.0523$ mas) differ by 5.5$\sigma$ and proper motions ($\mu_{\alpha} = -9.692\pm0.012$, $\mu_{\delta} = 7.349\pm0.012$ \masyr\, vs. $\mu_{\alpha} = -0.114\pm0.055$, $\mu_{\delta} = 6.419\pm0.053$ \masyr) differ by a factor of 2.
%
If the two stars are at the parallactic distance of \asas\ \citep[$d$ = 552.4 pc;][]{BailerJones21}, then their difference in proper motion ($\Delta\mu$ = $9.623\pm0.056$ \masyr) translates to a difference in tangential velocity of $\Delta V_{tan}$ = $25.37\pm0.21$ \kms.
%
The velocity difference is considerable given the projected separation of 2079\,au.
%
Using Kepler's 3rd law, and assuming the observed separation corresponds to the semi-major axis, and the observed $\Delta V_{tan}$ corresponded to the full orbital velocity, these quantities would predict a minimum system mass of $>150$ \msun.
%
Based on the implausible estimated dynamical mass inferred from the observed separation and difference in tangential velocities, we conclude that the visual companion 2MASS J08152298-3859244 is an interloper unrelated to \asas.

@@

The Remote Observatory Atacama Desert \citep[ROAD; ][]{Hambsch12} is a fully automated telescope located in Chile that obtains nightly photometry in multiple bands for a wide range of astronomical projects.
%
It consists of a 40-cm $f/6.8$ Optimized Dall-Kirkham and uses a Finger Lakes Instruments camera with a 4k$\times$4k array with pixels of $5\mu m$ in size.
%
Data are reduced using a custom pipeline and then published on the AAVSO website.

@@

The NEOWISE photometry is presented with the ASASSN $q$ light curve in Figure~\ref{fig:wisephot}.

@@

\subsection{LCOGT photometry}

RICHELLE TEXT HERE

@@

ATLAS is a project that searches for near earth asteroids down to a magnitude of 19 \citep{Tonry18}.
%
Two filters were obtained, the `o' and `c' filters respectively.
%
The data consists of two to four photometric points observed each night when conditions permitted.
%
Photometry with large errors was rejected in a first pass, then the remaining observations during a night were averaged and an error based on the r.m.s. of these nightly points was calculated.
%
The photometry covers the time period where the collision event occurred.

@@

A spectrum of the star was taken on 59829.8785 MJD equal to 2022-09-07 at 08:34:52 UTC from the Deep Sky Chile\footnote{\url{https://www.deepskychile.com/en/}} site in Chile (at longitude:70\degr 51\arcmin 11\farcs 86
latitude:30\degr S 31\arcmin 34\farcs 71 altitude:1700m AMSL).
%
The telescope is a Ritchey-Chretien telescope ($d=305$mm) at $f/5$ (with focal reducer CCD 67 astrophysics) on an equatorial mount GM 3000 HPS from 10 micron\footnote{\url{https://www.10micron.com/en/product/gm3000-hps/}}.
%
The spectrograph is a Spectrograph Alpy 600 $R=570$ with a $23\mu m$ wide slit\footnote{\url{https://www.shelyak.com/produit/alpy-600/?lang=en}} and an ATIK 414ex camera\footnote{\url{https://www.atik-cameras.com/product/atik-414ex/}}.
%
The spectrum is composed of three unbinned exposures of 1200s each, taken in automatic mode with Prism V11 Software\footnote{\url{https://www.prism-astro.com/}}, and the spectrum is processed with ISIS software\footnote{\url{http://www.astrosurf.com/buil/isis-software.html}}.
%
The final spectrum is shown in Figure~\ref{fig:2spotspectrum}.

@@

We measure the gradient of the light curve in an automated way by fitting straight lines to photometry between nights {\bf (XXX Richelle correct me on this!)}
%
We determine a robust lower limit to the transverse velocity of 5 km\,s$^{-1}$ for the material moving in front of the star, assuming a radius for the star of $1.07 R_\odot$.
%
This means that if the dust is on a circular orbit, it has to be within XXX au equivalent to an orbital period of YYY years around the star.

>>>>

The star 2MASS J08152329-3859234 underwent a sudden dimming event in December 2021, which was announced by \citet{RizzoSmith21} and assigned the identifier \asas, which we will subsequently use for the rest of this paper.
%
The star has undergone rapidly fluctuating photometry through to August 2022 \citep{RizzoSmith22} and has been monitored intensively by the AAVSO observers and the LCOGT network of telescopes.
%
It had previously shown no significant stellar variation in the optical bands for 2300 days as reported in \citet{RizzoSmith21}.
%
Searches through other photometric archives showed no significant changes in the optical bands before this epoch.
%
The wide field infrared satellite WISE has photometric imaging from 3.8 microns (band W1) through to 22 microns (band W4), and the NEOWISE survey has photometry for bands W1 and W2 for this star.
%
This star showed a significant brightening of 0.4 magnitudes in W1 and 0.8 magnitudes in W2 between two observing epochs (58055 MJD and 58245 MJD), and the IR color of the star had changed from W1-W2=0.0 to 1.2, but no significant changes in flux in the optical bands were seen during this time.
%
Some 900 days later, the dimming was seen in the optical, and the absorption is larger at shorter optical wavelengths.
%
These observations are all consistent with an event that generated a significant amount of sub-micron dust which subsequently started to transit the stellar disk.

@@

The properties of \asas\ are listed in Table~\ref{tab:Stellarprop}.
%
Photometry from various sources was compiled and analysed using the Virtual Observatory SED Analyzer \citep[VOSA; ][]{Bayo08} and the results are listed in Table~\ref{tab:Stellarprop}, showing that \asas\ is consistent with being a G2 type dwarf star.

@@

\subsection{Possible multiplicity of \asas} 
\label{subsec:companion}

\asas\ (Gaia DR3 5539970601632026752 = 2MASS~J08152329-3859234) has a neighbor (Gaia DR3 5539970597334497024 = 2MASS~J08152298-3859244) which is a visual double.
%
Based on the Gaia EDR3 mean ICRS position for epoch 2016.0, the visual companion lies at a separation $\rho$ = $3738.243\pm0.062$ mas and at position angle $\theta$ = $249^{\circ}.977$.
%
Their parallaxes ($\varpi$ = $1.7631\pm0.0112$ mas vs. $1.4711\pm0.0523$ mas) differ by 5.5$\sigma$ and proper motions ($\mu_{\alpha} = -9.692\pm0.012$, $\mu_{\delta} = 7.349\pm0.012$ \masyr\, vs. $\mu_{\alpha} = -0.114\pm0.055$, $\mu_{\delta} = 6.419\pm0.053$ \masyr) differ by a factor of 2.
%
If the two stars are at the parallactic distance of \asas\ \citep[$d$ = 552.4 pc;][]{BailerJones21}, then their difference in proper motion ($\Delta\mu$ = $9.623\pm0.056$ \masyr) translates to a difference in tangential velocity of $\Delta V_{tan}$ = $25.37\pm0.21$ \kms.
%
The velocity difference is considerable given the projected separation of 2079\,au.
%
Using Kepler's 3rd law, and assuming the observed separation corresponds to the semi-major axis, and the observed $\Delta V_{tan}$ corresponded to the full orbital velocity, these quantities would predict a minimum system mass of $>150$ \msun.
%
Based on the implausible estimated dynamical mass inferred from the observed separation and difference in tangential velocities, we conclude that the visual companion 2MASS J08152298-3859244 is an interloper unrelated to \asas.

@@

The Remote Observatory Atacama Desert \citep[ROAD; ][]{Hambsch12} is a fully automated telescope located in Chile that obtains nightly photometry in Astrodon B, V and I bands for a wide range of astronomical projects.
%
It consists of a 40-cm $f/6.8$ Optimized Dall-Kirkham and uses a Finger Lakes Instruments camera with a 4k$\times$4k array with pixels of $5\mu m$ in size.
%
Data are reduced using a custom pipeline and then published on the AAVSO website.

@@

The NEOWISE photometry is presented with the ASASSN $g$ light curve in Figure~\ref{fig:wisephot}.

@@

\subsection{LCOGT photometry}
Las Cumbres Observatory Global Telescope (LCOGT) is a network of 25 fully robotic operated telescopes distributed over 7 sites located all around the globe. These telescopes are designed to observe transient astronomical events at optical and near-infrared wavelengths. LCOGT provides a large variety of filter options, but the data we collected is in SDSS g, r and i bands. All data is automatically processed and calibrated by the BANZAI pipeline. This process involves bad-pixel masking, bias subtration, dark subtraction, flat field correction, source extraction, astrometric calibration and photometric calibration.

As mentioned in section \ref{subsec:companion}, \asas\ has a visual companion. This caused complications in BANZAIs automatic aperture extraction routine; sometimes correct apertures were extracted for both \asas\ and the nearby star, and sometimes both sources were extracted in one large aperture, often with an offset from the true center of \asas. Although the companion is faint, this caused the resulting photometry to be in inaccurate and therefor unusable. To correct this, the last two stages of the BANZAI routine, aperture extraction and photometry calibration, were modified for this specific situation.

\textbf{TODO: see commented notes}

% points to inlcude in text:
% lcogt observatory
% filters used (SDSS g, r, i)
% original pipeline
% visual companion, wrong source extraction
% adjust source extraction and photometry part of pipeline
% NEW PIPELINE:
% use sep package (cite) to subtract background from image, this is default / original extraction as used in banzai.
% replace apertures of asassn and visual companion
%   use high quality observation with good extraction as 'master frame' for reference
%   use ra 123.8470 and dec -38.9898 to find the corresponding center pixel using WCS (note that WCS info from astrometric calibration is in the header of the banzai pipeline result)
%   use ellipse parameters a b and theta of 'master frame'
%   use dynamic scaling, such that a_scaled, b_scaled = a*r_scale, b*r_scale where r_scale is the median aperture radius of the frame.
% compute the Kron radius of all sources (original kronrad computation)
% Replace the Kron radius of asassn-21qj by recomputing the Kron radius with an elliptical mask on the companion with a_kronmask = 5*a_comp, b_kronmask = 5*a_comp
% compute flux of all sources by summing over all the pixels in the ellipse scaled by 2.5 the kronradius (original flux computation)
% recompute the flux of asassn-21qj by masking out the companion, a_fluxmask = 3.8*a_comp, b_kronmask = 3.8*a_fluxmask, such that the pixels inside this mask do not count towards the flux of asassn-21qj.
% Call default banzai calibrate photometry to get calibrated magnitudes.

@@

ATLAS is a project that searches for near earth asteroids down to a magnitude of 19 \citep{Tonry18}.
%
Two filters were obtained, the `o' (orange) and `c' (cyan) filters respectively.
%
The data consists of two to four photometric points observed each night when conditions permitted.
%
Photometry with large errors was rejected in a first pass, then the remaining observations during a night were averaged and an error based on the r.m.s. of these nightly points was calculated.
%
The photometry covers the time period where the collision event occurred.

@@

A spectrum of the star was taken on 59829.8785 MJD, equal to 2022-09-07 at 08:34:52 UTC, from the Deep Sky Chile\footnote{\url{https://www.deepskychile.com/en/}} site in Chile (at longitude:70\degr 51\arcmin 11\farcs 86
latitude:30\degr S 31\arcmin 34\farcs 71 altitude:1700m AMSL).
%
The telescope is a Ritchey-Chretien telescope ($d=305$mm) at $f/5$ (with focal reducer CCD 67 astrophysics) on an equatorial mount GM 3000 HPS from 10 micron\footnote{\url{https://www.10micron.com/en/product/gm3000-hps/}}.
%
The spectrograph is a Spectrograph Alpy 600 $R=570$ with a $23\mu m$ wide slit\footnote{\url{https://www.shelyak.com/produit/alpy-600/?lang=en}} and an ATIK 414ex camera\footnote{\url{https://www.atik-cameras.com/product/atik-414ex/}}.
%
The spectrum is composed of three unbinned exposures of 1200s each, taken in automatic mode with Prism V11 Software\footnote{\url{https://www.prism-astro.com/}}, and the spectrum is processed with ISIS software\footnote{\url{http://www.astrosurf.com/buil/isis-software.html}}.
%
The final spectrum is shown in Figure~\ref{fig:2spotspectrum}.

@@

We measure the gradient of the light curve in an automated way by fitting straight lines to photometry between nights {\bf (XXX Richelle correct me on this!)}
%
We determine a robust lower limit to the transverse velocity of 5 km\,s$^{-1}$ for the material moving in front of the star, assuming a radius for the star of $1.009 R_\odot$.
%
This means that if the dust is on a circular orbit, it has to be within XXX au equivalent to an orbital period of YYY years around the star.
